# **ggplot2**

## Definitions
- geom: geometric object
- stats: statistical transformation
- aes: aesthetic attributes
- coord: coordinate system
- facet

```{r, "include" = FALSE}
library(ggplot2)
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Introduction to `ggplot2`: `qplot`

```{r}
data(diamonds); set.seed(1410)
d = diamonds[sample(nrow(diamonds), 100),]
print(paste(colnames(d), collapse = " "))
```

```{r}
qplot(log(carat), log(price), 
	data = d, color = color, 
	shape = cut, size = I(3), 
	alpha = I(0.7))
```


### Geom Objects
- point, smooth, boxplot, path, line, histogram, freqploy, density, bar


### smooth
```{r}
ggplot(d, aes(x = log(carat), y = log(price))) +
	geom_point() + 
	geom_smooth(
		se = TRUE, # standard error
		method = "loess", # default for small data size
		# "lm": linear, formula = y ~ poly(x, 2), polynomial
		# "rlm": insensitive to strange data point
		span = 0.7 # larger span, smoother line
		)
```

### jitter & boxplot
```{r}
ggplot(d, aes(x = color, y = price / carat)) + 
	geom_jitter(alpha = I(0.5)) # available: size, color, shape
ggplot(d, aes(x = color, y = price / carat)) + 
	geom_boxplot() # available: fill, size
``` 

### histogram & density
```{r}
ggplot(d, aes(x = carat)) + 
	geom_histogram(
		bins = 20, # number of bins, or using `binwidth`
		) +
	xlim(c(0, 3))
ggplot(d, aes(x = carat)) + 
	geom_density(aes(
		color = color, 
		fill = color, 
		alpha = I(0.7))) + 
	xlim(c(0, 3))
```

### bar
```{r}
ggplot(d, aes(x = color)) + 
	geom_bar(aes(
		weight = carat, # weighted by color
		color = color,
		fill = color))
```

### line & path
```{r}
data(economics); d = economics
ggplot(d, aes(x = date, y = unemploy / pop)) + geom_line()
year = function(x) as.POSIXlt(x)$year + 1900
ggplot(d, aes(x = unemploy / pop, y = uempmed, color = year(date))) + geom_path()
```

## `ggplot2` Grammar Components

### Layer

A layer consists of:
- Data and attributes mapping
- A statistical change
- A geometric object
- A position change

### Scale

Scale controls the mapping from data to aesthetic attributes. A scale is a function with some parameters.

- Size, shape, color ...

### Coordinate System

Scale change happens before statistical change, while coordinate change is after statistical change.

- Cartesian, semi-log, polar ...

### Data Structure

A graph object consists of data, mapping (default attributes mapping), layer, scale, coordinate system and facets.

## Build Graph with Layers

### Create a Graph Object

```{r}
d = diamonds[sample(nrow(diamonds), 100),]
graph = ggplot(d, aes(x = carat, y = price, color = cut))
graph
```

### Add Layer

Add `geom_...(aes, data, ..., stat, position)`
```{r}
graph = graph + geom_point()
graph
```

### Data Manipulation

- Change Data using `%+%` ...

```{r}
data(mtcars); d = mtcars
graph = ggplot(d, aes(x = mpg, y = wt, color = cyl)) + geom_point()
d$mpg = d$mpg ^ 2
graph %+% d
```

- or ...

```{r}
d = mtcars; graph = ggplot(d, aes(x = mpg, y = wt, colot = cyl)) + geom_point()
graph + geom_point(aes(x = mpg ^ 2))
```

- Remove data
`aes(x = NULL)`

- Grouped data

```{r}
library(nlme)
data(Oxboys); d = Oxboys
ggplot(d, aes(x = age, y = height)) + 
	geom_line(aes(color = Subject, group = Subject)) + 
	geom_smooth(aes(method = 'lm', group = 1), size = 2, se = FALSE, color = 'red')
```

### Statistical Change

- density

```{r}
d = diamonds[sample(nrow(diamonds), 100),]
ggplot(d, aes(carat)) + geom_histogram(aes(y = ..density..), binwidth = 0.1)
```

### Position Change

- dodge: side by side
- fill: normalize to 1
- identity: no change
- jitter
- stack

```{r}
graph = ggplot(d, aes(x = color, group = cut, color = cut, fill = cut))
g1 = graph + geom_bar(position = "dodge")
g2 = graph + geom_bar(position = "fill")
g3 = graph + geom_bar(position = "stack")
multiplot(g1, g2, g3, cols = 2)
```

## Practice

### Basic Graphic Type

```{r}
d = data.frame(x = c(3, 1, 5), y = c(2, 4, 6), label = LETTERS[1:3])
graph = ggplot(d, aes(x = x, y = y)) + xlab(NULL) + ylab(NULL)
g1 = graph + geom_point() + ggtitle('Point')
g2 = graph + geom_bar(stat = 'identity') + ggtitle('Bar')
g3 = graph + geom_line() + ggtitle('Line')
g4 = graph + geom_area() + ggtitle('Area')
g5 = graph + geom_path() + ggtitle('Path')
g6 = graph + geom_text(aes(label = label)) + ggtitle('Text')
g7 = graph + geom_tile() + ggtitle('Tile')
g8 = graph + geom_polygon() + ggtitle('Polygon')
multiplot(g1, g2, g3, g4, g5, g6, g7, g8, cols = 4)
```

## Scale, Axis and Legend

### Scale

```{r}
data(msleep); d = msleep
graph = ggplot(d, aes(x = sleep_total, y = sleep_cycle, color = vore)) + geom_point()
g2 = graph + scale_color_hue()
g3 = graph + scale_color_hue("Eat?", 
	breaks = c("herbi", "carni", "omni", "NA"), 
	labels = c("plants", "meat", "both", "don't know"))
g4 = graph + scale_color_brewer(palette = "Set2") # color set
multiplot(graph, g2, g3, g4, cols = 2)
```