# **ggplot2**

## Definitions
- geom: geometric object
- stats: statistical transformation
- aes: easthetic attributes
- coord: coordinate system
- facet

```{r, "include" = FALSE}
library(ggplot2)
```

## Introduction to `ggplot2`: `qplot`

```{r}
data(diamonds); set.seed(1410)
d = diamonds[sample(nrow(diamonds), 100),]
print(paste(colnames(d), collapse = " "))
```

```{r}
qplot(log(carat), log(price), 
	data = d, color = color, 
	shape = cut, size = I(3), 
	alpha = I(0.7))
```


### Geom Objects
- point, smooth, boxplot, path, line, histogram, freqploy, density, bar


### smooth
```{r}
ggplot(d, aes(x = log(carat), y = log(price))) +
	geom_point() + 
	geom_smooth(
		se = TRUE, # standard error
		method = "loess", # default for small data size
		# "lm": linear, formula = y ~ poly(x, 2), polynomial
		# "rlm": insensitive to strange data point
		span = 0.7 # larger span, smoother line
		)
```

### jitter & boxplot
```{r}
ggplot(d, aes(x = color, y = price / carat)) + 
	geom_jitter(alpha = I(0.5)) # available: size, color, shape
ggplot(d, aes(x = color, y = price / carat)) + 
	geom_boxplot() # available: fill, size
``` 

### histogram & density
```{r}
ggplot(d, aes(x = carat)) + 
	geom_histogram(
		bins = 20, # number of bins, or using `binwidth`
		) +
	xlim(c(0, 3))
ggplot(d, aes(x = carat)) + 
	geom_density(aes(
		color = color, 
		fill = color, 
		alpha = I(0.7))) + 
	xlim(c(0, 3))
```

### bar
```{r}
ggplot(d, aes(x = color)) + 
	geom_bar(aes(
		weight = carat, # weighted by color
		color = color,
		fill = color))
```

### line & path
```{r}
data(economics); d = economics
ggplot(d, aes(x = date, y = unemploy / pop)) + geom_line()
year = function(x) as.POSIXlt(x)$year + 1900
ggplot(d, aes(x = unemploy / pop, y = uempmed, color = year(date))) + geom_path()
```

## Build Graph with Layers

### Data

- Change Data using `%+%` ...

```{r}
data(mtcars); d = mtcars
graph = ggplot(d, aes(x = mpg, y = wt, color = cyl)) + geom_point()
d$mpg = d$mpg ^ 2
graph %+% d
```

- or ...

```{r}
d = mtcars; graph = ggplot(d, aes(x = mpg, y = wt, colot = cyl)) + geom_point()
graph + geom_point(aes(x = mpg ^ 2))
```

- Grouped data

```{r}
library(nlme)
data(Oxboys); d = Oxboys
ggplot(d, aes(x = age, y = height)) + 
	geom_line(aes(color = Subject, group = Subject)) + 
	geom_smooth(aes(method = 'lm', group = 1), size = 2, se = FALSE, color = 'red')
```

### Statistical Change

- density

```{r}
d = diamonds[sample(nrow(diamonds), 100),]
ggplot(d, aes(carat)) + geom_histogram(aes(y = ..density..), binwidth = 0.1)
```

### Position Change

- dodge: side by side
- fill: normalize to 1
- identity: no change
- jitter
- stack